---
title: "Project #1 - Data Visualization - Group 2"
output:
  word_document: default
  html_notebook: default
---

**Danielle do Val Goncalves Tudeia** - Student ID: 0823569

**Fernanda Barbieri de Camargo** - Student ID: 0825882

" The project represents our own work and we have adhered to St. Clair College's Academic Integrity policies in completing this project."

***\# R Version***

RStudio 2022.12.0+353 "Elsbeth Geranium" Release (7d165dcfc1b6d300eb247738db2c7076234f6ef0, 2022-12-03) for Windows

Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) RStudio/2022.12.0+353 Chrome/102.0.5005.167 Electron/19.1.3 Safari/537.36

    R version 4.2.2 (2022-10-31 ucrt)

***\# Import the tidyverse package because contains the ggplot2 used on data visualization***

```{r}
library(tidyverse)
```

***\# Check the tidyverse package version***

```{r}
packageVersion("tidyverse")
```

***\# Import the gganimate package to accomplish the data visualization's competition (we've already installed)***

```{r}
library(gganimate)
```

***\# Check the gganimate package version***

```{r}
packageVersion("gganimate")
```

***\# Information about the dataset used in this project***

[Name of data set:]{.underline} [Suicide Rates Overview 1985 to 2016](https://www.kaggle.com/datasets/russellyates88/suicide-rates-overview-1985-to-2016)

[Platform:]{.underline} Kaggle

[Creator's username:]{.underline} [Rusty](https://www.kaggle.com/russellyates88)

These data set is licensed by World Bank Dataset.

[Data set file's name]{.underline}: master.csv

***\# Changes done in the dataset before load it to R Studio***

The column ' I ' was almost totally empty so we removed it before load the file to R Studio. Now, there are 27,821 observations and 11 features on the dataset.

***\# Load the dataset's file to R Studio using the read.csv***

```{r}
suic = read.csv("C:/Users/Danielle/Documents/master.csv")
```

***\# Check the dataframe information such as features, data type, etc.***

```{r}
glimpse(suic)
```

```{r}
summary(suic)
```

***\# Visualize how the name of countries are written on the dataset***

```{r}
unique(suic$country)
```

***\# Define the country that is going to be used in the plots - We choose Guyana because the country has one of Top 5 suicide rate in the World. Then we create a subset only with Guyana's observations***

```{r}
suic_guyana <- suic %>% 
  filter(country == 'Guyana')
```

***\# Visualize the distinct years on the dataset for the country Guyana***

```{r}
unique(suic_guyana$year)
```

***\# Adjust the data on feature "Age", transforming "years" to "yrs" it will help the plot to be more "readable"***

```{r}
suic_guyana$age <- gsub("years", "yrs", suic_guyana$age)
```

***\# Create*** ***two plots displaying the distribution of a single continuous variable***

*\# To organize the data, create a new dataframe that has the mean of suicide rate aggregated by year*

```{r}
suic_year_rate_avg <- aggregate(suicides.100k.pop ~ year, suic_guyana, mean)
```

*\# Use ggplot to create a density plot: 'Distribution of Suicide Rate in Guyana'*

```{r}
suic_year_rate_avg %>% 
  ggplot() +
  geom_density(aes(x = suicides.100k.pop),
                 fill = 'pink',
                 colour = 'black')+
  labs(x = 'Suicides (per 100k Population)') +
  labs(y = 'Density')+
  labs(subtitle = 'from 1988 to 2013')+
  labs(caption = "Note: From 1988 to 2013, the anual suicide rate average \n was more frequent between 27 and 33 suicides/100k population")+
  ggtitle('Distribution of Suicide Rate in Guyana')+
  theme_bw()
```

*\# Use ggplot to create the histogram: 'Distribution of Suicide Rate in Guyana'*

```{r}
suic_year_rate_avg %>% 
  ggplot() +
  geom_histogram(aes(x=suicides.100k.pop),
                 fill = 'pink',
                 colour = 'black',
                 binwidth = 5)+
  labs(x = 'Suicides (per 100k population)')+
  labs(y = 'Number of Years')+
  labs(subtitle = 'from 1988 to 2013')+
  labs(caption = "Note: From 1988 to 2013, the anual suicide rate average \n stayed between 27 and 33 suicides/100k population for 8 years")+
  ggtitle('Distribution of Suicide Rate in Guyana')+
  theme_bw()
```

***\# Create*** ***two plots displaying the distribution of a single categorical variable***

*\# To organize the data, create a new dataframe that has the sum of suicides number aggregated by age*

```{r}
suic_age_sum <- aggregate(suicides_no ~ age, suic_guyana, sum)
```

*\# Use ggplot to create a plot: 'Suicides in Guyana per Age'*

```{r}
suic_age_sum %>% 
  ggplot(aes(x = fct_relevel(age,'5-14 yrs','15-24 yrs','25-34 yrs','35-54 yrs','55-74 yrs','75+ yrs'), y = suicides_no)) +
  geom_col(fill = 'pink',
           color = 'black')+
  labs(x = 'Age') +
  labs(y = 'Number of Suicides')+
  labs(subtitle = 'from 1988 to 2013')+
  labs(caption = "Note: From 1988 to 2013, the number of suicides was more \n frequent among people who had between 35 and 54 years old")+
  ggtitle('Suicides in Guyana per Age')+
  theme_bw()
```

*\# To organize the data, create a new dataframe that has the sum of suicides number aggregated by gender*

```{r}
suic_sex_sum <- aggregate(suicides_no ~ sex, suic_guyana, sum)
```

*\# Use ggplot to create a plot: 'Suicides in Guyana per Gender'*

```{r}
suic_sex_sum %>% 
  ggplot(aes(x = sex, y = suicides_no)) +
  geom_col(fill = 'pink',
           color = 'black')+
  labs(x = 'Gender') +
  labs(y = 'Number of Suicides')+
  labs(subtitle = 'from 1988 to 2013')+
  labs(caption = "Note: From 1988 to 2013, men commit more than twice as many suicides as women")+
  ggtitle('Suicides in Guyana per Gender')+
  theme_bw()
```

***\# Create*** ***one plot displaying information about both a continuous variable and a categorical variable***

*\# To organize the data, create a new dataframe that has the mean of suicide rate aggregated by age and year*

```{r}
suic_age_rate_avg <- aggregate(suicides.100k.pop ~ age + year, suic_guyana, mean)
```

*\# Use ggplot to create a boxplot: 'Suicide Rate in Guyana per Age'*

```{r}
suic_age_rate_avg %>% 
  ggplot() +
  geom_boxplot(aes(x = fct_relevel(age,'5-14 yrs','15-24 yrs','25-34 yrs','35-54 yrs','55-74 yrs','75+ yrs'), y = suicides.100k.pop),
               fill = 'pink')+
  labs(x = 'Age') +
  labs(y = 'Suicides (per 100k Population)')+
  labs(subtitle = 'from 1985 to 2016')+
  labs(caption = "Note: On average, from 1988 to 2013, people between \n the ages of 24 and 35 years old have the highest suicide rate")+
  ggtitle('Suicide Rate in Guyana per Age')+
  theme_bw()
```

***\# Create two plots should display information that shows a relationship between two variables***

*\# Filter the dataframe to leave only the people who probably is economic active (\>=15 and \<75 years)*

```{r}
suic_guyana_prod <- suic_guyana %>% 
  filter(suic_guyana$age != '5-14 yrs' & suic_guyana$age != '75+ yrs')
```

*\# To organize the data, create a new dataframe that has the average of suicide rate aggregate by gdp per capita*

```{r}
suic_pop_year_avg <- aggregate(suicides.100k.pop ~ gdp_per_capita...., suic_guyana_prod, mean)
```

*\# Use ggplot to create the plot: 'Suicide Rate vs GDP per Capita'*

```{r}
suic_pop_year_avg %>% 
  ggplot(aes(x=gdp_per_capita...., y=suicides.100k.pop)) +
  geom_point()+
  geom_smooth(fill = 'pink',
              colour = 'black',
              linewidth = 0.5)+
  labs(x = 'GDP per Capita ($)') +
  labs(y = 'Suicides (per 100k population)')+
  labs(subtitle = 'Age between 15 and 75 yrs')+
  labs(caption = "Note: The GDP per capita doesn't have correlation to the Suicide Rate")+
  ggtitle('Suicide Rate vs GDP per Capita')+
  theme_bw()
```

*\# To organize the data, create a new dataframe that has the sum of population aggregate by year*

```{r}
suic_year_pop_sum <- aggregate(population ~ year, suic_guyana, sum)
```

*\# Create a new dataframe with the feature population (sum) and suicide rate (average)*

```{r}
suic_pop_rate <- suic_year_pop_sum %>% 
  add_column(suicides.100k.pop = suic_year_rate_avg$suicides.100k.pop)

```

*\# Use ggplot to create the plot: 'Suicide Rate vs Population'*

```{r}
suic_pop_rate %>% 
  ggplot(aes(x=population, y=suicides.100k.pop)) +
  geom_point()+
  geom_smooth(fill = 'pink',
              colour = 'black',
              linewidth = 0.5)+
  labs(x = 'Population') +
  labs(y = 'Suicides (per 100k population)')+
  labs(subtitle = 'from 1988 to 2013')+
  labs(caption = "Note: The population doesn't have correlation to the Suicide Rate")+
  ggtitle('Suicide Rate vs Population')+
  theme_bw()
```

***\# Create one plot use faceting and displaying information about 4 variables***

*\# To organize the data, create a new dataframe that has the sum of suicide rate aggregate by year, sex and age*

```{r}
sum_suic_sex_age <- aggregate(suicides.100k.pop ~ sex + generation + year, suic_guyana, mean)
```

*\# Use ggplot to create the plot 'Suicide rate in Guyana over the Years split by Sex and Generation'*

```{r}
sum_suic_sex_age %>% 
  filter(year >=2000) %>% 
  ggplot() +
  geom_line(aes(x = year, y = suicides.100k.pop),
            size = 0.5,
            colour = 'blue') +
  facet_grid(sex ~ generation) +
  geom_point(aes(x = year, y = suicides.100k.pop),
             colour = 'blue')+
  labs(x = 'Year') +
  labs(y = 'Suicides (per 100k population)')+
  labs(subtitle = 'by Sex and Generation')+
  labs(caption = "Note: The highest suicide rates are for men from Generation X, however,\n there is a growing trend among Millenials. Generation Z has the lower rates")+
  ggtitle('Suicide rate in Guyana over the Years')+
  theme_bw()
```

***\# Create competition plot: an opportunity to explore what's possible and get creative***

*\# Create a new dataframe only containing information about men between 15 and 24 years old*

```{r}
suic_guyana_comp <- suic_guyana %>% 
  filter(suic_guyana$sex =='male' & year >=2000)
```

*\# To organize the data, create a new dataframe that has the sum of suicide rate aggregate by year, sex and age*

```{r}
suic_guyana_comp_avg <- aggregate(suicides.100k.pop ~ generation + year,suic_guyana_comp, mean)
```

*\# Use ggplot and gganimate to create an animated plot: 'Male suicide rate in Guyana over the Years'*

```{r}
suic_plt_comp <- suic_guyana_comp_avg %>% 
  ggplot() +
  geom_line(aes(x = year, y = suicides.100k.pop, color = generation),
            linewidth = 0.8) +
  labs(x = 'Year') +
  labs(y = 'Suicides (per 100k population)')+
  labs(color = 'Generation')+
  labs(subtitle = 'by Generation')+
  labs(caption = "Note: The highest suicide rates are for men from Generation X, however,\n there is a growing trend among Millenials")+
  ggtitle('Male suicide rate in Guyana over the Years')+
  theme_bw()

```

```{r}
suic_plt_comp +
  transition_reveal(year)
```

<http://localhost:10749/session/file716814fe4e4d.gif>
